Dispersion distance and the matter distribution of the Universe in dispersion space

### 2015-06-04 writeup/dm_space_distortions.tex
rewording beyond the scope to future work
slight rewordings.

Large-scale structure is usually studied through its two-point statistics, most
commonly the power spectrum, $P(k)$.
Unlike the redshift-space distortions, Equation~\ref{e:delta_s} does not have a
simple form in harmonic space. The equation's third term couples harmonic modes
and thus the two-point statistics cannot be phrased as a simple power spectrum.
We will instead use $C^{ss}_\ell(\chi,\chi')$, which is
the cross-correlation angular power spectrum of the dispersion-space over-density,
on shells at $\chi$ and $\chi'$:
\begin{align}
    &C^{ss}_\ell(\chi, \chi') =
    \int\ud\Omega\ud\Omega' Y_{\ell m}(\hat n) Y^*_{\ell m}(\hat n')
    \langle \delta_{fs}(\hat n \chi) \delta_{fs}(\hat n' \chi')
        \rangle.
\end{align}
The first two terms in Equation~\ref{e:delta_s} are stationary. For stationary,
isotropic, tracers $x$ and $y$ we have 
$\langle \delta_x(\vec k, \chi) \delta_y(\vec k', \chi) \rangle = (2\pi)^3
\delta^3(\vec k - \vec k') P_{xy}(k, \chi)$.  If for a moment we ignore
structure evolution, the angular cross-power spectrum of such tracers is
\begin{align}
    %\delta_{\ell\ell'}\delta_{mm'}C^{xy}_\ell(\chi, \chi')
    %&= \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
    %    \langle \delta_x(\hat n \chi) \delta_y(\hat n' \chi') \rangle,
    %    \\
    C^{xy}_\ell(\chi,\chi')
    &= \frac{2}{\pi}
\int_0^\infty\ud k k^2 j_\ell(k\chi) j_{\ell}(k\chi')P_{xy}(k).
\end{align}
In reality the power spectrum evolves on the order of a Hubble time.
The angular cross correlations will be very small unless $\chi$ and $\chi'$ are
within a few correlation lengths of one another, roughly a hundred
megaparsecs.  The evolution of the power spectrum is negligible over these time 
differences which leads to a straight-forward way to include the evolution of
the perturbations:
\begin{align}
    % I've tuned the alignment of the domain (using &)
    % here to get the equation number and
    % domain on the same line. Not elegant.
    C^{xy}_\ell(\chi,\chi') \approx&
        \frac{2}{\pi}
        \int_0^\infty\ud k k^2
        j_\ell(k\chi) j_{\ell}(k\chi')
        P_{xy}(k,(\chi + \chi')/2),\nonumber \\
    &|\chi - \chi'| \ll 1/aH.
\end{align}

@@

Perhaps, more concerning is the dispersion from the source's immediate environment
as well as the halo in which it resides. These electrons will presumably be
clustered around the source on scales far smaller than the survey
resolution and are thus not accounted for in the
$\delta_e$--$\delta_f$
correlations included in the above formalism.  This will have a mean
contribution, common to all sources and which may be epoch dependant, that will
modify the mean dispersion--distance relation. This will need to be calibrated in some
way or else fit with nuisance parameters during parameter estimation.  There
will also be a stochastic piece which varies from source to source.  This will
limit the precision at which $\chi_s$ can be measured, thus limiting the
resolution of dispersion space density maps in the radial direction.

Detailed treatment of the above complications is beyond the scope of this
work. The magnitude of the contribution from the progenitor is at present
unknown. To crudely deal with it here we use relatively large
bins widths of $\Delta\chi = 100\,{\rm Mpc}/h$
and exclude the $\chi = \chi'$ bin pairs.
Because the local term drops rapidly with radial separations, this effectively 
discards the contribution from the local term which is most sensitive to having
precise radial distances.

>>>>

Large-scale structure is usually studied through its two-point statistics, most
commonly the power spectrum, $P(k)$.
Unlike the redshift-space distortions, Equation~\ref{e:delta_s} does not have a
simple form in harmonic space. The equation's third term couples harmonic modes
and thus the two-point statistics cannot be phrased as a simple power spectrum.
We will instead use $C^{ss}_\ell(\chi,\chi')$, which is
the cross-correlation angular power spectrum of the dispersion-space over-density,
on shells at $\chi$ and $\chi'$:
\begin{align}
    &C^{ss}_\ell(\chi, \chi') =
    \int\ud\Omega\ud\Omega' Y_{\ell m}(\hat n) Y^*_{\ell m}(\hat n')
    \langle \delta_{fs}(\hat n \chi) \delta_{fs}(\hat n' \chi')
        \rangle.
\end{align}
The first two terms in Equation~\ref{e:delta_s} are stationary. For stationary,
isotropic, tracers $x$ and $y$ we have 
$\langle \delta_x(\vec k, \chi) \delta_y(\vec k', \chi) \rangle = (2\pi)^3
\delta^3(\vec k - \vec k') P_{xy}(k, \chi)$.  If for a moment we ignore
structure evolution, the angular cross-power spectrum of such tracers is
\begin{align}
    %\delta_{\ell\ell'}\delta_{mm'}C^{xy}_\ell(\chi, \chi')
    %&= \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
    %    \langle \delta_x(\hat n \chi) \delta_y(\hat n' \chi') \rangle,
    %    \\
    C^{xy}_\ell(\chi,\chi')
    &= \frac{2}{\pi}
\int_0^\infty\ud k k^2 j_\ell(k\chi) j_{\ell}(k\chi')P_{xy}(k).
\end{align}
In reality the power spectrum evolves on the order of a Hubble time.
The angular cross correlations will be very small unless $\chi$ and $\chi'$ are
within a few correlation lengths of one another, roughly a hundred
megaparsecs.  The evolution of the power spectrum is negligible over these time
differences which leads to a straight-forward way to include the evolution:
\begin{align}
    % I've tuned the alignment of the domain (using &)
    % here to get the equation number and
    % domain on the same line. Not elegant.
    C^{xy}_\ell(\chi,\chi') \approx&
        \frac{2}{\pi}
        \int_0^\infty\ud k k^2
        j_\ell(k\chi) j_{\ell}(k\chi')
        P_{xy}(k,(\chi + \chi')/2),\nonumber \\
    &|\chi - \chi'| \ll 1/aH.
\end{align}

@@

Perhaps, more concerning is the dispersion from the source's immediate environment
as well as the halo in which it resides. These electrons will presumably be
clustered around the source on scales far smaller than the survey
resolution and are thus not accounted for in the
$\delta_e$--$\delta_f$
correlations included in the above formalism.  This will have a mean
contribution, common to all sources and which may be epoch dependant, that will
modify the mean dispersion--distance relation. This will need to be calibrated in some
way or else fit with nuisance parameters during parameter estimation.
There will also be a stochastic piece which varies from source to source.  This will
limit the precision at which $\chi_s$ can be measured, thus limiting the
resolution of dispersion space density maps in the radial direction.
We note that one-point statistics from the same dataset 
should provide empirical information about these properties independent
of their physical origin.

Detailed treatment of the above source modelling is beyond the scope of this
work. The magnitude of the contribution from the progenitor is at present
unknown. To crudely deal with it here we use relatively large
bins widths of $\Delta\chi = 100\,{\rm Mpc}/h$
and exclude the $\chi = \chi'$ bin pairs.
Because the local term drops rapidly with radial separations, this effectively 
discards the contribution from the local term which is most sensitive to having
precise radial distances.
